---
title: "Introduction to fishmetricR package"
author: 
- "Zárate-Salazar, J. Rafael, PhD"
- "Researcher | PPEC - UFS"
- "rzatatesalazar@gmail.com"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to fishmetricR package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Introduction

The **fishmetricR** package was developed to facilitate obtaining allometric coefficients of fish from the FishBase database [fishbase](https://fishbase.se/).

This vignette presents a simple example of how to use the package's main function to extract the coefficients of a species' length-weight equation.

# Install

You can install the package directly from GitHub:

```r
# install.packages("remotes")
remotes::install_github("jrz-s/fishmetricR")
```
# Application example

```{r}

library(fishmetricR)

# Example 1: Retrieve coefficients for a unique species
output1 <- fishmetricR::get_fish_allometric(
                          genus = "Brycon"
                         ,species = "orthotaenia")

print(output1)

## Export with kable function
knitr::kable(output1
             ,caption = "Get coefficients for a unique species")

# Example 2: Retrieve coefficients for multiple species
output2 <- fishmetricR::get_fish_allometric(
                          genus = c("Brycon","Clarias","Piaractus")
                         ,species = c("orthotaenia","gariepinus","brachypomus"))

print(output2)

## Export with kable function
knitr::kable(output2
             ,caption = "Get coefficients for multiple species")

```

# Function details

The get_allometric() function performs web scraping of FishBase to collect the allometric coefficients.

The data returned includes:

- \( a \): slope of the length-weight equation

- \( b \): exponent of the length-weight equation

It also includes other relevant information such as minimum and maximum sizes, location, sample size, R², etc.

## Allometric equation

FishBase provides length-weight relationships based on the classical allometric equation:

$$
W = a \cdot L^b
$$

Where:

- \( W \) is the body weight (usually in grams),
- \( L \) is the total, standard, or fork length (usually in centimeters),
- \( a \) is the intercept coefficient (related to body form), and
- \( b \) is the allometric exponent (related to growth pattern).

To estimate these coefficients using linear regression, the equation is log-transformed into its linear form:

$$
\log_{10}(W) = \log_{10}(a) + b \cdot \log_{10}(L)
$$

This transformation allows the estimation of \( \log_{10}(a) \) and \( b \) using ordinary least squares. FishBase provides the values of \( a \) and \( b \) obtained through this procedure. The parameter \( b \) is biologically informative — for example:

- \( b \approx 3 \): isometric growth (body shape remains proportional),
- \( b > 3 \): positive allometric growth (fish become stouter as they grow),
- \( b < 3 \): negative allometric growth (fish become more elongated as they grow).

# References

- Froese, R. & Pauly, D. (Eds). FishBase. World Wide Web electronic publication. Acess: https://fishbase.se

- **fishmetricR** package documentation.
